---
title: "fars"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{fars}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, echo = FALSE, include = FALSE}
library(CourseraBuildingRPckg)
library(dplyr)
library(maps)
```

*CourseraBuildingRPckg is a peer reviewed project to practice R package Building*  

This packages provides functions to **process** and **plot** data from the US National Highway Traffic Safety Administration's [Fatality Analysis Reporting System](https://www.nhtsa.gov/research-data/fatality-analysis-reporting-system-fars).


* Process data with `fars_summarize_years()`

* Visualize data with `fars_map_state()`

Those functions **require** that current working directory contains FARS data from the given year named like **accident_[year].csv.bz2**


## The data

The packages has 3 files relating the FARS data from 2013 to 2015.

They are accessible under the folder accessible by running the following line.

```{r,eval=FALSE}
system.file("extdata", package = "CourseraBuildingRPckg")
```

Those files can also be used by setting current directory as following right before calling the function (in the same chunk):

```{r}
setwd(system.file("extdata", package = "CourseraBuildingRPckg"))
```




## Summarizing data 

Let's process the accidents data of 2013 by summarizing them by month in year in a tibble using the `fars_summarize_years()` function.


```{r}
setwd(system.file("extdata", package = "CourseraBuildingRPckg")) 
#Needed if you want to use data from the package, otherwise function will look for files in the directory the function is run.
getwd()
fars_2013 <- fars_summarize_years(2013)
fars_2013
```

Nota Bene `fars_summarize_years()` can also create a summary report for multiple years.

```{r}
setwd(system.file("extdata", package = "CourseraBuildingRPckg"))
fars_2013_2014 <- fars_summarize_years(c(2013, 2014))
fars_2013_2014
```

## Visualizing data

The `fars_map_state` function will plot all accidents locations for a given year and federal state on a US federal state map.

```{r}
setwd(system.file("extdata", package = "CourseraBuildingRPckg"))
fars_map_state(26, 2014)
```

